---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
steps <- read.csv("activity.csv")
#steps$time <- strptime( sprintf("%04d", steps$interval), "%H%M" ) 
```


## What is mean total number of steps taken per day?
1.  total number of steps taken per day
```{r}
stepsByDate <- split(steps$steps, steps$date)
sumByDate <- sapply(stepsByDate, sum, USE.NAMES = TRUE, na.rm=TRUE)
sumByDate

```

2.  A histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(sumByDate, bin = 2000, main="Total Number of Steps Taken Each Day", xlab="Total steps per Day")
```

3.  The mean and median of the total number of steps taken per day
```{r}
totalStepsMean <- mean(sumByDate)
totalStepsMean
totalStepsMedian <- median(sumByDate)
totalStepsMedian
```

## What is the average daily activity pattern?
1.  A time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 
```{r}
stepsByInt <- split(steps$steps, steps$interval)
meanByInt <- sapply(stepsByInt, mean, USE.NAMES = TRUE, na.rm=TRUE)
meantime <- as.numeric(names(meanByInt))
plot(meantime, meanByInt, type="l", main="Average Daily Activity Pattern", xlab="time", ylab="average steps taken")
```

2.  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxInd <- which.max(meanByInt)
names(maxInd)
```


## Imputing missing values





## Are there differences in activity patterns between weekdays and weekends?
