---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
steps <- read.csv("activity.csv")
steps$date <- as.Date(steps$date)
#steps$time <- strptime( sprintf("%04d", steps$interval), "%H%M" ) 
```


## What is mean total number of steps taken per day?
1.  total number of steps taken per day
```{r}
stepsByDate <- split(steps$steps, steps$date)
sumByDate <- sapply(stepsByDate, sum, USE.NAMES = TRUE, na.rm=TRUE)
sumByDate

```

2.  A histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(sumByDate, bin = 2000, main="Total Number of Steps Taken Each Day", xlab="Total steps per Day")
```

3.  The mean and median of the total number of steps taken per day
```{r}
totalStepsMean <- mean(sumByDate)
totalStepsMean
totalStepsMedian <- median(sumByDate)
totalStepsMedian
```

## What is the average daily activity pattern?
1.  A time series plot of the 5-minute interval and the average number of steps taken, averaged across all days 
```{r}
stepsByInt <- split(steps$steps, steps$interval)
meanByInt <- sapply(stepsByInt, mean, USE.NAMES = TRUE, na.rm=TRUE)
meantime <- as.numeric(names(meanByInt))
plot(meantime, meanByInt, type="l", main="Average Daily Activity Pattern", xlab="time", ylab="average steps taken")
dev.off()
```

2.  Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxInd <- which.max(meanByInt)
names(maxInd)
```


## Imputing missing values
1.  The total number of missing values in the dataset
```{r}
sum(is.na(steps$steps))
```

2.   A new dataset with all of the missing values filled in using the mean for that 5-minute interval.
```{r}
stepsNaInd <- which(is.na(steps$steps))
stepsNoNA <- steps
for (i in stepsNaInd){
    curTime <- stepsNoNA$interval[i]
    avgAtInt <- meanByInt[as.character(curTime)]
    stepsNoNA$steps[i] <- avgAtInt
}
```

3.  New sum for the imputed dataset 
```{r}
stepsByDateNoNA <- split(stepsNoNA$steps, stepsNoNA$date)
sumByDateNoNA <- sapply(stepsByDateNoNA, sum, USE.NAMES = TRUE)
```

4.  Summary of the Imputed data set

* A histogram of the total number of steps taken each day
```{r}
qplot(sumByDateNoNA, bin = 2000, main="Total Number of Steps Taken Each Day (NA Imputed)", xlab="Total steps per Day")
```

* the mean and median total number of steps taken per day
```{r}
totalStepsMeanNoNA <- mean(sumByDateNoNA)
totalStepsMeanNoNA
totalStepsMedianNoNA <- median(sumByDateNoNA)
totalStepsMedianNoNA
```

* comparing with the original data set:  
  After imputing the dataset, both the mean and the median increase.  This is expected because the missing values were treated as 0 before, but are now replaced with its correspoinding 5-minute-interval average.  

## Are there differences in activity patterns between weekdays and weekends?
1.  Creating a new factor variable in the dataset indicating whether a given date is a weekday or weekend day.
```{r}
for (i in 1:length(steps$date)){
    steps$wday[i] <- if( weekdays(steps$date[i]) == "Saturday" || weekdays(steps$date[i]) == "Sunday" ) "weekend" else "weekday"
}
steps$wday <- as.factor(steps$wday)
```

2.  Time series plots of the average number of steps taken between weekdays and weekends
```{r}
library(dplyr)
stepMeans <- steps %>% group_by(wday, interval) %>% summarise(steps=mean(steps, na.rm=TRUE))
library(lattice)
xyplot(steps ~ interval| wday, data = stepMeans,type = "l",xlab = "Interval", ylab = "Avg num of steps", main="Activity Patterns between Weekdays and Weekends", layout=c(1,2))
```